Bootstrapping Domain Specific Languages for Neurally-Guided Bayesian Program Learning

### 2018-02-08 paper.tex
jo\begin{figure}

How should we perform inference for the model
in~\ref{graphicalModel}(a)?
We take inspiration from the Helmholtz machine and the Wake/Sleep algorithm~\cite{hinton1995wake},
which alternates between ``waking'', which draws samples from a recognition model (our $q$);
and  ``sleeping'', which trains a recognition model on samples from the generative model (our $\mathcal{D},\theta$).
For us, a wake cycle means using the recognition model to search for programs that solve the tasks.
The wake cycle can be thought of as a kind of active problem solving,
guided by the current generative and recognition models.
A sleep cycle means updating these models so that the DSL and  recognition model
more closely match the distribution of tasks.
This wake/sleep style of inference is the reason why we call our approach \system.

@@

Our goal  is to induce a DSL and find good programs solving each of the tasks.
Our strategy is to iterate through three steps: (1) a \emph{wake} phase where we search for programs, and then a pair of \emph{sleep} phases: (2) learning a neural recognition model -- which we use
to accelerate the search over programs -- and (3) improving the DSL.
The key observation here is that these three steps bootstrap off each other (Figure~\ref{feeding}):

\noindent \textbf{Searching for programs:}  Our program search  is informed by both the DSL and the recognition model.
When these improve we find better programs.

\noindent\textbf{Learning a recognition model:} The recognition model is trained both on samples from the DSL and on programs found by the search procedure. As the DSL improves and as search finds more programs, the recognition model gets both more data to train on and better data.

\noindent\textbf{Improving the DSL:} We induce a DSL from the programs we have found so far which solve the tasks;
  as we solve more tasks, we can hone in on richer DSLs that more closely match the domain.

@@

    \draw [->,thick] (synthesis.-120) to[out = -150,in = 60] node[below,rotate = 45,align = center]{{\footnotesize Trains}\\{\footnotesize (Sleep)}} (DSL.30);
    \draw [->,thick] (synthesis.-60) to[out = -30,in = 120] node[below,rotate=-45,align = center]{{\footnotesize Trains}\\{\footnotesize (Sleep)}} (recognitionModel.150);
    \draw [->,thick] (DSL.east) to[out = -30,in = 210] node[above, align = center]{{\footnotesize Trains}\\{\footnotesize (Sleep)}} (recognitionModel.west);

@@

@@

\subsection{Sleep Phase I: Learning a Recognition Model}\label{recognitionSection}

@@

\subsection{Sleep Phase II: Inducing the DSL}\label{grammarInductionSection}

>>>>

How should we perform inference for the model
in~\ref{graphicalModel}(a)?
We take inspiration from the Helmholtz machine and the Wake/Sleep algorithm~\cite{hinton1995wake},
which alternates between ``waking'', which draws samples from a recognition model (our $q$);
and  ``sleeping'', which trains a recognition model on samples from the generative model (our $\mathcal{D},\theta$).
For us, a wake cycle means using the recognition model to search for programs that solve the tasks.
The wake cycle can be thought of as a kind of active problem solving,
guided by the current generative and recognition models.
A sleep cycle means  updating these models so that the DSL and  recognition model
more closely match the distribution of tasks.
An important difference from the Helmholtz machine
is that our setup has two distinct sleep phases: a phase where we update the generative model
$(\mathcal{D},\theta)$, which we call \textbf{G-Sleep},
and a second phase where we update the recognition model $q$,
which we call \textbf{R-Sleep}.
This wake/sleep style of inference is the reason why we call our approach \system.

@@

Our goal  is to induce a DSL and find good programs solving each of the tasks.
Our strategy is to iterate through three steps: (1) searching for programs; (2) learning a neural recognition model -- which we use
to accelerate the search over programs -- and (3) improving the DSL.
The key observation here is that these three steps bootstrap off each other (Figure~\ref{feeding}):
\\\noindent \textbf{Wake: Searching for programs.}  Our program search  is informed by both the DSL and the recognition model.
When these improve we find better programs.
\\\noindent\textbf{R-Sleep: Learning a recognition model.} The recognition model is trained both on samples from the DSL and on programs found by the search procedure. As the DSL improves and as search finds more programs, the recognition model gets both more data to train on and better data.
\\\noindent\textbf{G-Sleep: Improving the DSL.} We induce a DSL from the programs we have found so far which solve the tasks;
  as we solve more tasks, we can hone in on richer DSLs that more closely match the domain.

@@

    \draw [->,thick] (synthesis.-120) to[out = -150,in = 60] node[below,rotate = 45,align = center]{{\footnotesize Trains}\\{\footnotesize (G-Sleep)}} (DSL.30);
    \draw [->,thick] (synthesis.-60) to[out = -30,in = 120] node[below,rotate=-45,align = center]{{\footnotesize Trains}\\{\footnotesize (R-Sleep)}} (recognitionModel.150);
    \draw [->,thick] (DSL.east) to[out = -30,in = 210] node[above, align = center]{{\footnotesize Trains}\\{\footnotesize (R-Sleep)}} (recognitionModel.west);

@@

\begin{figure}
\begin{tikzpicture}

  \begin{scope}[shift = {(0.7,0)}]  
    \node[obs] at (3,3) (dx){$\mathcal{D}$};
  \node[latent] at (3.5,1.75) (zp){$p$};
  \node[obs] at (4,3) (tx){$\theta^{(x)}$};
  \node[obs] at (3.5,0.5) (xp) {$x$};
  \node[align = center] at ([yshift = -0.6cm,xshift = 0.5cm]xp.south) {Wake: Infer $p$};
  \draw [->,red] (xp.east) to[out = 0,in = 20] node(nn){} (tx.north);
  \draw [->,red] (tx.south) -- (zp.north);
  \draw [->] (dx.south) -- (zp.north);

  \node at (nn) {
    \begin{tikzpicture}[x=2.5cm,y=1.25cm,transform canvas={scale=0.2,shift={+(-1,2.5)}}]
      \tikzstyle{neuron}=[circle,fill=blue!50,minimum size=20pt]
      \fill[fill=white] (-0.25,-0.5) rectangle (2.25,-4.5);
      \node[rectangle,width=2cm] at (1,1) {};
      \foreach \name / \y in {1,...,4}
          \node[neuron] (I-\name) at (0,-\y) {};
      \foreach \name / \y in {1,...,3}
          \node[neuron] (H-\name) at (1,-\y-0.5) {};
      \foreach \name / \y in {1,...,4}
          \node[neuron] (O-\name) at (2,-\y) {};
      \foreach \source in {1,...,4}
          \foreach \dest in {1,...,3}
              \draw [-latex] (I-\source) -- (H-\dest);
      \foreach \source in {1,...,3}
          \foreach \dest in {1,...,4}
              \draw [-latex] (H-\source) -- (O-\dest);
    \end{tikzpicture}
  };
  \node[shift={+(0,-0.55)}] at (nn) { $q(x)$ };
  \node[shift={+(0,-0.55)}] at (nn) { $q(x)$ };
  \node[shift={+(0,-0.55)}] at (nn) { $q(x)$ };

    \begin{scope}[shift = {(5,0)}]  
      \node[obs] at (3,3) (dt){$\mathcal{D},\theta$};
      \node[obs] at ([yshift = -0.7cm,xshift = 0.0cm]dt.south) (p2){$p_2$};
      \node[obs] at ([yshift = -0.7cm,xshift = 0.0cm]p2.south) (x2){$x_2$};
      \draw [->] (p2.south) -- (x2.north);
      \draw [->] (dt.south) -- (p2.north);

      \node[obs] at ([yshift = 0cm,xshift = -0.5cm]p2.west) (p1){$p_1$};
      \node[obs] at ([yshift = -0.7cm,xshift = 0.0cm]p1.south) (x1){$x_1$};
      \draw [->] (p1.south) -- (x1.north);
      \draw [->] (dt.south) -- (p1.north);

      \node[obs] at ([yshift = 0cm,xshift = 0.5cm]p2.east) (p3){$p_3$};
      \node[obs] at ([yshift = -0.7cm,xshift = 0.0cm]p3.south) (x3){$x_3$};
      \draw [->] (p3.south) -- (x3.north);
      \draw [->] (dt.south) -- (p3.north);

      \node at ([yshift = 0cm,xshift = 0.3cm]p3.east) {$\cdots $};
      \node at ([yshift = 0cm,xshift = 0.3cm]x3.east) {$\cdots $};

  \node[obs] at ([yshift = 0cm,xshift = 1.3cm]p3.east) (zp){$p$};
  \node[obs] at ([yshift = 0cm,xshift = 1.3cm]x3.east) (xp) {$x$};
  \draw [->] (zp.south) -- (xp.north);
  \plate {}{(zp)(xp)}{$x\in X$};

  \draw[dashed,cyan,very thick] ([yshift = 0.5cm,xshift = -0.5]p1.west) rectangle  ([yshift = -0.5cm,xshift = -0.5]xp.east);
  \node[align = center] at ([yshift = -1cm,xshift = 0.6cm]x2.south) {R-Sleep: Train $q$\\training data: cyan $(x,p)$};
    \end{scope}

    \begin{scope}[shift = {(5,-6)}]  
      \node[latent] at (0.5,3) (d){$\mathcal{D}$};
  \node[latent] at (1.5,3) (t){$\theta$};
  \node[obs] at (1,1.75) (z){$p$};
%  \node[latent] at (3.5,1.5) (tx){$\theta^{(x)}$};
  \node[obs] at (1,0.5) (x) {$x$};
  \edge {z}{x};
  \edge {d,t}{z};
  \plate {}{(z)(x)}{$x\in X$};
  \node[align = center] at ([yshift = -1cm]x.south) {G-Sleep: Induce $(\mathcal{D},\theta})$};
    \end{scope}

  \end{tikzpicture}
  \end{figure}

@@

\subsection{R-Sleep: Learning a Recognition Model}\label{recognitionSection}

@@

\subsection{G-Sleep: Inducing the DSL}\label{grammarInductionSection}
